Forecasting the 2024 U.S. Presidential Election

### 2024-10-31 paper/paper.qmd

## Overview

@@

## Measurement and Limitations

The primary limitations of this analysis stem from the quality and variability of polling data, which can introduce biases through factors like sample size and pollster methodology. Additionally, the model provides a static snapshot of voter preferences, without accounting for the dynamic nature of elections For instance, shifts in public opinion over time or the impact of campaign events. The exclusion of key external factors, such as economic conditions or voter turnout, also limits the model's ability to fully capture the complexities of election outcomes. Furthermore, the simplified Electoral College simulation assumes that polling data will accurately predict state-level results, which may not always be the case, especially in battleground states with volatile voter behavior.

## Outcome variables
Our primary outcome variable is the percentage of support for Kamala Harris and Donald Trump in each poll. This variable is central to our analysis because it directly captures voter preference, which is the most relevant metric for forecasting election outcomes. By modeling the percentage of support, we can quantify how various factors—such as pollster quality, sample size, and state-level dynamics—influence the candidates' standings in the polls. This outcome variable also allows us to simulate the Electoral College results by aggregating predicted support across different states, which is crucial for determining the likelihood of either candidate winning the presidency.

@fig-support below shows the distribution of support for both candidates across the polls included in the dataset.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig.pos: "H"
#| label: fig-support
#| fig-cap: Distribution of Support for Kamala Harris and Donald Trump

ggplot(data, aes(x = pct, fill = candidate_name)) +
  geom_histogram(bins = 30, alpha = 0.7, position = "identity") +
  facet_wrap(~ candidate_name) +
  labs(x = "Percentage of Support",
       y = "Count") +
  scale_fill_discrete(name = "Candidates") +
  theme_minimal() +
  theme(legend.position = "bottom")

@@

>>>>

## Overview
Our study utilizes polling data from FiveThirtyEight's 2024 Presidential Election Forecast Database [@FiveThirtyEightPollingData2024], a comprehensive polling dataset maintained by ABC news. This database compiles and standardizes polling results from various organizations, applying quality metrics and assessments to each poll. Several key variables from this dataset are crucial to our analysis:

@@

## Measurement and Limitations
There are several measurement and limitation considerations for our dataset:

- **Poll Quality:** While we filter for high-quality polls using numeric grades, differences in polling methodologies may introduce systematic biases. The inclusion of pollster ratings helps account for historical accuracy but cannot completely eliminate these potential biases.
- **Temporal Dynamics:** Our dataset provides a snapshot of voter preferences during a specific timeframe. This static nature means we cannot capture the full dynamics of voter preference evolution over the campaign period.
- **Geographic Coverage:** Although we have national and state-level polling data, coverage varies by state. Battleground states typically have more frequent polling, while safer states may have sparse data, potentially affecting our state-level predictions.
- **Response Bias:** Despite careful methodology by pollsters, self-selection bias in survey participation and social desirability bias in responses remain potential concerns.

## Outcome variables
Our primary outcome variable is the percentage of support (*pct*) for Kamala Harris and Donald Trump in each poll. This measurement represents the proportion of respondents who indicate they would vote for each candidate if the election were held on the day of the poll. The variable directly captures voter preferences and serves as the foundation for our electoral predictions.

@fig-support below illustrates the distribution of support for both candidates, revealing several notable patterns. First, the support percentages cluster between 40% and 60%, reflecting the competitive nature of the race. Second, the distributions show slight differences between candidates, with Harris's support displaying more variation than Trump's. This pattern might reflect differences in voter certainty or polling methodology across different states and time periods.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: fig-support
#| fig-cap: Distribution of Support for Kamala Harris and Donald Trump

ggplot(data, aes(x = pct, fill = candidate_name)) +
  geom_histogram(bins = 30, alpha = 0.8, position = "identity") +
  facet_wrap(~ candidate_name) +
  labs(x = "Percentage of Support",
       y = "Count") +
  scale_fill_discrete(name = "Candidates") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

# Appendix {#sec-appendix}
